---
title: "Week 6 Assignment"
author: "Carmen Hoyt"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

# Connect database

## Load Packages

```{r}
library(tidyverse)
library(DBI)
library(duckdb)
library(dbplyr)
```

## Open Connection

```{r}
conn <- DBI::dbConnect(duckdb::duckdb(), dbdir="database/database.db")

# Look at tables
DBI::dbListTables(conn)
```

# Who were the winners?

Who measured the most eggs?

```{r}
# Load tables
egg_table <- tbl(conn, "Bird_eggs")
egg_table

bird_nests <- tbl(conn, "Bird_nests")
bird_nests

personnel <- tbl(conn, "Personnel")
personnel

# Query
egg_table %>%
group_by(Nest_ID) %>%
  summarise(measured_eggs = max(Egg_num, na.rm = TRUE)) %>%
  left_join(bird_nests, by = "Nest_ID") %>%
  group_by(Observer) %>%
  summarise(total_eggs = sum(measured_eggs, na.rm = TRUE)) %>%
  left_join(personnel, by = c("Observer" = "Abbreviation")) %>%
  select(Name, total_eggs) %>%
  slice_max(order_by = total_eggs, n = 3) %>%
  show_query()
```

Vanessa Loverti won. 

dbplyr used one query.

It limited to the first three rows using an R expression `(q04 <= 3)`.

```{r}
DBI::dbDisconnect(conn, shutdown = TRUE)
```


# Characterising egg variation

## Step 1

```{r}
# Load tables
egg_table <- tbl(conn, "Bird_eggs")
egg_table

bird_nests <- tbl(conn, "Bird_nests")
bird_nests

species <- tbl(conn, "Species")
species

# Query
egg_table %>%
  distinct(Nest_ID) %>%
  left_join(bird_nests, by = "Nest_ID") %>%
  left_join(species, by = c("Species" = "Code")) %>%
  distinct(Species, Scientific_name) %>%
  arrange(Scientific_name) %>%
  compute(name = "temp_species_table")
```

## Step 2

```{r}
species_query = """SELECT Code, Scientific_name FROM temp_species_table"""
for row in c.execute(species_query).fetchall():  # DuckDB lame-o workaround
    species_code = row[0]
    scientific_name = row[1]
    # query egg data for that species (step 3)
    # compute statistics and print results (step 4)
```

